---
title: "AR siRNA Knockdown of GFP/Puncta Count/Nuclei Number of Objects"
author: "Abby Coleman"
date: "`r Sys.Date()`"
geometry: margin = 1cm
output: pdf_document
---

```{r setup, include=FALSE}
# devtools::install_github("jkaupp/nord")
# # CRAN version
# install.packages("nord")
library(nord)
library(here)
library(pander)
library(ggplot2)
library(ggbeeswarm)
library(tidyverse)
library(wesanderson)
library(kableExtra)
library(viridis)
source(here("functions.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r read_in_data}
df <- readRDS(here("data", "cleaned", "by_nucleus.RDS"))
```

# General Conclusions for "2024-10-03" experiment

ken: can you give us total puncta per well as an output?


# Raw Signal by Condition for 1003

```{r raw_bp_1003, fig.height = 8, fig.width = 10, warning = FALSE, results = "asis"}
for_plot <- df %>%
  filter(compound != "PC3 GFP-AR K22") %>%
  filter(exp_date == "2024-10-03") %>%
  mutate(nice_compound2 = case_when(ar_sirna_dose_ul == 0 ~ nice_compound,
                                    TRUE ~ "siRNA Conditions")) %>%
  ungroup()
#list raw (unnormalized) variables to plot
var_list <- c("average_puncta_nuc", "avg_gfp_nuc", "nuclei_number_of_objects")
#run plot_raw_bp (defined in functions.R) to plot each raw variable by condition
bps <- lapply(var_list, plot_raw_bp_1003, for_plot)
```

notes:

* no difference in nuclei number of objects between days-- biological/technical issue
* ken: can you give us total puncta per well as an output?

\newpage

```{r mystery, results = "asis", include = FALSE}
mysteries <- for_plot %>%
  ungroup() %>%
  filter(nuclei_number_of_objects < 8 | average_puncta_nuc < 2 | avg_gfp_nuc > 512) %>%
  dplyr::select(exp_date, treatment_duration, compound, row, column, average_puncta_nuc, avg_gfp_nuc, nuclei_number_of_objects)
names(mysteries) <- c("Exp\nDate", "Treatment\nDuration", "Compound", "Row", "Column", "Avg\nPuncta", "Avg\nGFP", "Nuclei\nCount")
pandoc.table(mysteries, caption = "Outliers", keep.line.breaks = TRUE)
```

```{r stat_tab1, results = "asis"}
for_cv_tab <- df %>%
  ungroup() %>%
  filter(exp_date == "2024-10-03") %>%
  dplyr::select(c("exp_date", "treatment_duration", names(df)[grepl("plate_cv", names(df))])) %>%
  distinct()
pandoc.table(for_cv_tab, caption = "Plate %CV")
```


```{r stat_tab2, results = "asis"}
for_zprime_tab <- df %>% 
  ungroup() %>%
  filter(exp_date == "2024-10-03") %>%
  dplyr::select(c("treatment_duration", names(df)[grepl("z_prime", names(df))])) %>%
  dplyr::select(-c("z_prime_puncta", "z_prime_gfp", "z_prime_nuc")) %>%
  distinct()
names(for_zprime_tab) <- c("Treatment\nDuration", "Puncta\nReagent", "Puncta\nScramble", "Puncta\nNo Treatment", "GFP\nReagent", "GFP\nScramble", "GFP\nNo Treatment", "Nucleus Count\nReagent", "Nucleus Count\nScramble", "Nucleus Count\nNo Treatment")
pandoc.table(for_zprime_tab, keep.line.breaks = TRUE, caption = "Z'Factor")

```

\newpage

# General Conclusions for "2024-08-14" and "2024-09-05" experiments

  * There doesn't seem to be a large enough separation between the high and low controls in the GFP output. Will have to see Z'Factor calculations to be totally sure. 
  * $R^2$ of .547 in normalized puncta vs GFP linear model-- good but not great correlation between the two metrics. Also I don't totally trust the normalized GFP because of the control separation.


# Raw Signal by Condition, first two experiments

```{r raw_bp, fig.width = 8, warning = FALSE, results = "asis", out.width = "75%"}
for_plot <- df %>%
  filter(compound != "PC3 GFP-AR K22") %>%
  filter(exp_date %in% c("2024-08-14", "2024-09-05")) %>%
  mutate(nice_compound2 = case_when(ar_sirna_dose_ul == 0 ~ nice_compound,
                                    TRUE ~ "siRNA Conditions"))
#list raw (unnormalized) variables to plot
var_list <- c("average_puncta_nuc", "avg_gfp_nuc")
#run plot_raw_bp (defined in functions.R) to plot each raw variable by condition
bps <- lapply(var_list, plot_raw_bp, for_plot)
```

  * Separation between high and low controls for GFP seems very small.
  * Might want to think about using a different reagent concentration for low control? Or setting controls for each dose of reagent?

\newpage

# Normalized Signal by Condition

```{r norm_bp, warning = FALSE, fig.width = 10, results = "asis", out.width = "75%"}
for_plot <- df %>%
  filter(exp_date %in% c("2024-08-14", "2024-09-05")) %>%
  filter(compound != "PC3 GFP-AR K22") %>%
  mutate(nice_compound2 = case_when(ar_sirna_dose_ul == 0 ~ nice_compound,
                                    TRUE ~ "siRNA Conditions"))
#list raw (unnormalized) variables to plot
var_list <- c("puncta_percent_control", "gfp_percent_control")
#run plot_raw_bp (defined in functions.R) to plot each raw variable by condition
bps <- lapply(var_list, plot_norm_bp, for_plot)
```

Big difference between +/-R1881 in normalized puncta count, not so much in GFP.

\newpage

# Puncta vs GFP

```{r nn_v_ctg_scatterplot, warning = FALSE, fig.width = 8}
for_plot <- df %>%
  filter(exp_date %in% c("2024-08-14", "2024-09-05")) %>%
  filter(compound != "PC3 GFP-AR K22") %>%
  filter(exp_date == "2024-08-14")
#linear model
linmod <- lm(gfp_percent_control ~ puncta_percent_control, data = for_plot)
linmod_summ <- summary(linmod)
coef <-  round(linmod_summ$coefficients[2,1], 3)
pval <-  linmod_summ$coefficients[2,4]
rsq <- round(linmod_summ$r.squared, 3)
if (pval < .001) {
  pval <- formatC(pval, format = "e", digits = 3)
} else {
  pval <- round(pval, 3)
}
#plot
scatterplot <- ggplot(data = for_plot, aes(x = puncta_percent_control, y = gfp_percent_control, color = nice_compound_nobreak)) +
  geom_point(aes(shape = r1881), size = 3, alpha = .5) +
  theme_bw() +
  ggtitle("Puncta vs GFP") +
  xlab("Percent Control Puncta Count") +
  ylab("Percent Control GFP") +
  geom_smooth(color = "cornflowerblue", method = "lm") +
  scale_color_viridis_d(name = "Condition") +
  scale_shape_discrete(name = "R1881") +
  xlim(c(-100, 150)) +
  ylim(c(-100, 150)) +
  theme(text=element_text(family="serif"))

# +
#   geom_label(aes(x = max(for_plot$puncta_percent_control, na.rm = TRUE), y = min(for_plot$gfp_percent_control, na.rm = TRUE)), label = paste("Pval:", pval), hjust = 1)
#   geom_label(aes(x = max(for_plot$puncta_percent_control, na.rm = TRUE), y = min(for_plot$gfp_percent_control, na.rm = TRUE) + 30), label = paste("Rsq:", rsq), hjust = 1) +
#   theme(text=element_text(family="serif")) +
#   geom_label(aes(x = max(for_plot$puncta_percent_control, na.rm = TRUE), y = min(for_plot$gfp_percent_control, na.rm = TRUE) + 60), label = paste("Coef:", coef), hjust = 1) +
#   xlim(c(min(for_plot$puncta_percent_control, na.rm = TRUE), max(for_plot$puncta_percent_control, na.rm = TRUE))) +
#   ylim(c(min(for_plot$gfp_percent_control, na.rm = TRUE), max(for_plot$gfp_percent_control, na.rm = TRUE)))

scatterplot
```


 
  * Weird GFP controls make their normalized values all over the place-- note differences of scale in GFP vs puncta count.
  * Reagent doses kind of muddy any conclusions we can try and make about dose-effectiveness. Why were there so many different levels of reagents tested in these experiments?

## Linear Model Coefficients:
  * $R^2$ = `r rsq`
  * Model coefficient = `r coef`
  * p-value = `r pval`

## Percent Control Formulas:

$$
Percent Control Puncta Count = \frac{Avg Number of Puncta Per Nucleus - Plate Average Puncta Low Control}{Plate Average Puncta High Control - Plate Average Puncta Low Control}*100
$$
$$
Percent Control GFP = \frac{Average GFP Intensity Per Nucleus - Plate Average GFP Low Control}{Plate Average GFP High Control - Plate Average GFP Low Control}*100
$$


